---
output:
  pdf_document: 
    fig_caption: yes
  html_document: default
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 50
---

############################################## 

# ISLRv2, 3.7, Q-15

### Boston data set, per capita crime rate is the response, and the other variables are the predictors.

### (a) For each predictor, fit a simple linear regression model to predict the response. Describe your results. In which of the models is there a statistically significant association between the predictor and the response? Create some plots to back up your assertions

```{r warning=FALSE}
library(MASS)
library(ISLR2)
```

```{r}
# explore structure of Boston dataset
str(Boston)
```

### per capita crime rate is the response,

```{r}
# change colname
df_1 <- Boston
colnames(df_1)[colnames(df_1) == "crim"] <- "response"
```

\newpage

```{r}
# run lm models using single-predictors
lm_boston_1 <- lm(data = df_1, formula = response ~ df_1$zn)
lm_boston_2 <- lm(data = df_1, formula = response ~ df_1$indus)
lm_boston_3 <- lm(data = df_1, formula = response ~ df_1$chas)
lm_boston_4 <- lm(data = df_1, formula = response ~ df_1$nox)
lm_boston_5 <- lm(data = df_1, formula = response ~ df_1$rm)
lm_boston_6 <- lm(data = df_1, formula = response ~ df_1$age)
lm_boston_7 <- lm(data = df_1, formula = response ~ df_1$dis)
lm_boston_8 <- lm(data = df_1, formula = response ~ df_1$rad)
lm_boston_9 <- lm(data = df_1, formula = response ~ df_1$tax)
lm_boston_10 <- lm(data = df_1, formula = response ~ df_1$ptratio)
lm_boston_11 <- lm(data = df_1, formula = response ~ df_1$lstat)
lm_boston_12 <- lm(data = df_1, formula = response ~ df_1$medv)
```

\newpage

```{r}
# Extract p-values
df_p_values <- data.frame('model_name'=NA,'predictor'=NA,'p_value'=0.0)
```

```{r}
new_row <- c(
          'model_name' = 'lm_boston_1',
          'predictor' = 'zn',
          'p_value' = summary(eval(lm_boston_1))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_2',
          'predictor' = 'indus',
          'p_value' = summary(eval(lm_boston_2))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_3',
          'predictor' = 'chas',
          'p_value' = summary(eval(lm_boston_3))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_4',
          'predictor' = 'nox',
          'p_value' = summary(eval(lm_boston_4))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_5',
          'predictor' = 'rm',
          'p_value' = summary(eval(lm_boston_5))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_6',
          'predictor' = 'age',
          'p_value' = summary(eval(lm_boston_6))$coefficients[2, 4]
          )
```

\newpage

```{r}
df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_7',
          'predictor' = 'dis',
          'p_value' = summary(eval(lm_boston_7))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_8',
          'predictor' = 'rad',
          'p_value' = summary(eval(lm_boston_8))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_9',
          'predictor' = 'tax',
          'p_value' = summary(eval(lm_boston_9))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_10',
          'predictor' = 'ptratio',
          'p_value' = summary(eval(lm_boston_10))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_11',
          'predictor' = 'lstat',
          'p_value' = summary(eval(lm_boston_11))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_12',
          'predictor' = 'medv',
          'p_value' = summary(eval(lm_boston_12))$coefficients[2, 4]
          )

df_p_values <- rbind(df_p_values, new_row)
```

\newpage

```{r}
df_p_values
```

#### 'chas' predictor has p_value \> 0.05, so not significant.

#### Rest of the predictors have p_value \< 0.05, so significant.

\newpage

```{r}
par(mfrow = c(1, 1))
plot(df_1)
```

\newpage

```{r}
plot(y=df_1$response, x=df_1$chas)
```

#### from above plots, 'response' and 'chas' do NOT exhibit a linear relationship.

\newpage

## (b) Fit a multiple regression model to predict the response using all of the predictors. Describe your results. For which predictors can we reject the null hypothesis H0 : b1 = 0?

```{r}
# Fit multiple regression model
lm_boston_all <- lm(data = df_1, formula = response ~ .)
summary(lm_boston_all)
```

#### based on multiple linear regression model using ALL predictors,

#### p-values\<0.05 for (zn, dis, rad, medv), so significant, so can reject null hytothesis H0 : b1 = 0

#### p-values\>0.05 for (indus, chas, nox, rm, age,tax, ptratio, lstat), so NOT significant, so CANNOT reject null hytothesis H0 : b1 = 0

\newpage

## (c) How do your results from (a) compare to your results from (b)?

Create a plot displaying the univariate regression
coefficients from (a) on the x-axis, and the
multiple regression coefficients from (b) on the
y-axis. That is, each predictor is displayed as a
single point in the plot. Its coefficient in a
simple linear regression model is shown on the
x-axis, and its coefficient estimate in the
multiple linear regression model is shown on the
y-axis.

```{r}
# store univariate coefficients
univar_zn <- summary(eval(lm_boston_1))$coefficients[2, 1]
univar_indus <- summary(eval(lm_boston_2))$coefficients[2, 1]
univar_chas <- summary(eval(lm_boston_3))$coefficients[2, 1]
univar_nox <- summary(eval(lm_boston_4))$coefficients[2, 1]
univar_rm <- summary(eval(lm_boston_5))$coefficients[2, 1]
univar_age <- summary(eval(lm_boston_6))$coefficients[2, 1]
univar_dis <- summary(eval(lm_boston_7))$coefficients[2, 1]
univar_rad <- summary(eval(lm_boston_8))$coefficients[2, 1]
univar_tax <- summary(eval(lm_boston_9))$coefficients[2, 1]
univar_ptratio <- summary(eval(lm_boston_10))$coefficients[2, 1]
univar_lstat <- summary(eval(lm_boston_11))$coefficients[2, 1]
univar_medv <- summary(eval(lm_boston_12))$coefficients[2, 1]
```

```{r}
df_coeff <- data.frame(
    'zn' = univar_zn,
    'indus' = univar_indus,
    'chas'= univar_chas,
    'nox'= univar_nox,
    'rm'= univar_rm,
    'age'= univar_age,
    'dis'= univar_dis,
    'rad'= univar_rad,
    'tax'= univar_tax,
    'ptratio' = univar_ptratio,
    'lstat' = univar_lstat,
    'medv' = univar_medv
)
rownames(df_coeff) <- c("uni_variate")
```

\newpage

```{r}

df_multivar_coeff <- data.frame(
    'zn' = summary(eval(lm_boston_all))$coefficients[-1,1][['zn']],
    'indus' = summary(eval(lm_boston_all))$coefficients[-1,1][['indus']],
    'chas'= summary(eval(lm_boston_all))$coefficients[-1,1][['chas']],
    'nox'= summary(eval(lm_boston_all))$coefficients[-1,1][['nox']],
    'rm'= summary(eval(lm_boston_all))$coefficients[-1,1][['rm']],
    'age'= summary(eval(lm_boston_all))$coefficients[-1,1][['age']],
    'dis'= summary(eval(lm_boston_all))$coefficients[-1,1][['dis']],
    'rad'= summary(eval(lm_boston_all))$coefficients[-1,1][['rad']],
    'tax'= summary(eval(lm_boston_all))$coefficients[-1,1][['tax']],
    'ptratio' = summary(eval(lm_boston_all))$coefficients[-1,1][['ptratio']],
    'lstat' = summary(eval(lm_boston_all))$coefficients[-1,1][['lstat']],
    'medv' = summary(eval(lm_boston_all))$coefficients[-1,1][['medv']]
)
rownames(df_multivar_coeff) <- c("multi_variate")
```

```{r}
# multi_variate row
df_coeff <- rbind(df_coeff, df_multivar_coeff[1,])
```

```{r message=FALSE, warning=FALSE}
library(data.table)
# transpose to get rownames as predictors
df_coeff_transpose <- transpose(df_coeff)
rownames(df_coeff_transpose) <- colnames(df_coeff)
colnames(df_coeff_transpose) <- rownames(df_coeff)
df_coeff_transpose
```

#### The coefficient estimates for predictors show changes.

('indus', 'nox', 'age', 'tax', 'ptratio') above
list has a negative coefficient in multi_variate,
had a positive coeff. in uni_variate regression.
('zn','rm',) above list has a positive coefficient
in multi_variate, had a negative coeff. in
uni_variate regression.

\newpage

```{r warning=FALSE}
plot(x=df_coeff_transpose$uni_variate, y = df_coeff_transpose$multi_variate,
     xlab = 'uni_variate', ylab= "multi_variate", main = "multi_variate_coeff V uni_variate_coeff")
```

\newpage

## (d) Is there evidence of non-linear association between any of the predictors and the response? To answer this question, for each predictor X, fit a model of the form Y = b0 + b1\*X + b2\**X\**X + b3\**X\**X\*X + error.

```{r}
# lm_boston_poly_3 <- lm(data = df_1, formula = response ~ poly(df_1$chas,3))
# Error in poly(df_1$chas, 3) : 'degree' must be less than number of unique points
# due to error, do not use model, do not use predictor for poly regression.
```

```{r}
# run polynomial models (cubic) for single-predictors
lm_boston_poly_1 <- lm(data = df_1, formula = response ~ poly(df_1$zn,3))
lm_boston_poly_2 <- lm(data = df_1, formula = response ~ poly(df_1$indus,3))
# lm_boston_poly_3 results in error, so not using.
lm_boston_poly_4 <- lm(data = df_1, formula = response ~ poly(df_1$nox,3))
lm_boston_poly_5 <- lm(data = df_1, formula = response ~ poly(df_1$rm,3))
lm_boston_poly_6 <- lm(data = df_1, formula = response ~ poly(df_1$age,3))
lm_boston_poly_7 <- lm(data = df_1, formula = response ~ poly(df_1$dis,3))
lm_boston_poly_8 <- lm(data = df_1, formula = response ~ poly(df_1$rad,3))
lm_boston_poly_9 <- lm(data = df_1, formula = response ~ poly(df_1$tax,3))
lm_boston_poly_10 <- lm(data = df_1, formula = response ~ poly(df_1$ptratio,3))
lm_boston_poly_11 <- lm(data = df_1, formula = response ~ poly(df_1$lstat,3))
lm_boston_poly_12 <- lm(data = df_1, formula = response ~ poly(df_1$medv,3))
```

```{r}
df_poly_p_values <- data.frame('model_name'=NA,'predictor'=NA,
                               'p_value_x_pow_1'=0.0, 'p_value_x_pow_2'=0.0, 'p_value_x_pow_3'=0.0)
```

\newpage

```{r}
# store p_values.
# FUTURE IMPROVEMENT: use functions and/or loops
new_row <- c(
          'model_name' = 'lm_boston_poly_1',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_1))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_1))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_1))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_2',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_2))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_2))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_2))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

# new_row <- c(
#           'model_name' = 'lm_boston_poly_3',
#           'predictor' = 'zn',
#           'p_value_x_pow_1' = NA,
#           'p_value_x_pow_2' = NA,
#           'p_value_x_pow_3' = NA
#           )

# df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_4',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_4))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_4))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_4))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_5',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_5))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_5))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_5))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)
```

\newpage

```{r}


new_row <- c(
          'model_name' = 'lm_boston_poly_6',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_6))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_6))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_6))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_7',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_7))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_7))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_7))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_8',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_8))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_8))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_8))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_9',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_9))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_9))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_9))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_10',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_10))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_10))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_10))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_11',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_11))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_11))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_11))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)

new_row <- c(
          'model_name' = 'lm_boston_poly_12',
          'predictor' = 'zn',
          'p_value_x_pow_1' = summary(eval(lm_boston_poly_12))$coefficients[2, 4],
          'p_value_x_pow_2' = summary(eval(lm_boston_poly_12))$coefficients[3, 4],
          'p_value_x_pow_3' = summary(eval(lm_boston_poly_12))$coefficients[4, 4]
          )

df_poly_p_values <- rbind(df_poly_p_values, new_row)
```

\newpage

```{r}
# recast to numeric type from character type
# round for easier display and interpretation
df_poly_p_values$p_value_x_pow_1 <- round(as.numeric(df_poly_p_values$p_value_x_pow_1),3)
df_poly_p_values$p_value_x_pow_2 <- round(as.numeric(df_poly_p_values$p_value_x_pow_2),3)
df_poly_p_values$p_value_x_pow_3 <- round(as.numeric(df_poly_p_values$p_value_x_pow_3),3)
```

```{r}
alpha_threshold = 0.05 # threshold for significance of p-value

#binary test against threshold for significance of p-value
df_poly_p_values$p_value_x_pow_1_test <- ifelse(df_poly_p_values$p_value_x_pow_1 <= alpha_threshold,1,0) 
df_poly_p_values$p_value_x_pow_2_test <- ifelse(df_poly_p_values$p_value_x_pow_2 <= alpha_threshold,1,0) 
df_poly_p_values$p_value_x_pow_3_test <- ifelse(df_poly_p_values$p_value_x_pow_3 <= alpha_threshold,1,0) 

```

```{r}
df_poly_p_values_test <- subset(df_poly_p_values[ ,c(1,6:8)])

rownames(df_poly_p_values_test)<- rownames(df_poly_p_values)
```

\newpage

```{r}
df_poly_p_values_test <- df_poly_p_values_test[-1,]
df_poly_p_values_test
```

#### based on above output, for p-value significance test at threshold \<=0.05, resulting in (1 as \<=0.05) and (0 as \>0.05) for all poly models, x_pow_1 (x raised to 1) survives p-value test for all poly models, x_pow_2 (x raised to 2) survives p-value test. 

#### THIS IS EVIDENCE OF NON-LINEAR RELATIONSHIP BETWEEN RESPONSE AND PREDICTORS. 

#### for certain poly models, x_pow_3 (x raised to 3) survives p-value test, indicating that in certain 1-predictor cases for this dataset the cubic_fit results in better fit to training data (in this case ALL data is training data since no train-test split).
