---
output:
  pdf_document: 
    fig_caption: yes
  html_document: default
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 50
---

############################################## 

# ISLRv2, 3.7, Q-14
```{r}
set.seed (1)
x1 <- runif (100)
x2 <- 0.5 * x1 + rnorm (100) / 10
y <- 2 + 2 * x1 + 0.3 * x2 + rnorm (100)
```

## The last line corresponds to creating a linear model in which y is
a function of x1 and x2. Write out the form of the linear model.
What are the regression coefficients?
```{r}
# form of linear model y <-2+ 2.15x1 +error_new
# error_new is combination of errors resulting from combining terms
# coefficients x1 = 2
# coefficients x2 = 0.3
```

## (b) What is the correlation between x1 and x2? Create a scatterplot
displaying the relationship between the variables.
```{r}
cor(x1,x2)
```


```{r}
plot(x1,x2)
```

## (c) Using this data, fit a least squares regression to predict y using
x1 and x2. Describe the results obtained. 
```{r}
lm_1 <- lm(y ~ x1 + x2)
summary(lm_1)
```
#### Results displayed above, F-stat is relatively low, still higher than 1; 
associated p-value <0.05 so relationship exists between predictors and response.
p-value for x2 is NOT <0.05, so x2 not significant predictor.
p-value for x1 is <0.05, so x1 is a significant predictor. 
Although this p-value is very close to 0.05.


###  What are b0_hat, b1_hat, and b2_hat? 
```{r}
# b0_hat = 2.1305
# b1_hat = 1.4396
# b2_hat = 1.0097
```

How do these relate to the true b0, b1, and b2? 
```{r}
# y <- 2 + 2 * x1 + 0.3 * x2 + rnorm (100) # original formula

# b0_hat = 2.1305 > b0= 2
# b1_hat = 1.4396 < b1= 2
# b2_hat = 1.0097 > b2= 0.3
```

Can you reject the null hypothesis H0 : b1 = 0? 
```{r}
# p-value for x1 is <0.05, so x1 is a significant predictor. 
#Although this p-value is very close to 0.05. 
# so can reject null hypothesis H0 : b1 = 0
```

How about the null hypothesis H0 : b2 = 0?
```{r}
# p-value for x2 is NOT<0.05, so x2 is NOT a significant predictor. 
# so cannot reject null hypothesis H0 : b2 = 0
```

## (d) Now fit a least squares regression to predict y using only x1. Comment on your results. 
```{r}
lm_2 <- lm(y ~ x1)
summary(lm_2)
```

#### Results displayed above, F-stat is increased from previous run, 
still higher than 1; associated p-value <0.05, 
so relationship exists between predictors and response.
p-value for x1 is <0.05, so x1 is a significant predictor. 
Although this p-value is very close to 0 and very far away from 0.05.

## Can you reject the null hypothesis H0 : b1 = 0?
```{r}
# p-value for x1 is <<0.05, so x1 is a significant predictor. 
# so can reject null hypothesis H0 : b1 = 0
```

## (e) Now fit a least squares regression to predict y using only x2. Comment on your results.
```{r}
lm_3 <- lm(y ~ x2)
summary(lm_3)
```
#### Results displayed above, F-stat is decreased from previous run, 
still higher than 1; associated p-value <0.05, 
so relationship exists between predictors and response.
p-value for x2 is <0.05, so x2 is a significant predictor. 
Although this p-value is very close to 0 and very far away from 0.05.

## Can you reject the null hypothesis H0 : b1 = 0?
```{r}
# p-value for x2 is <<0.05, so x2 is a significant predictor. 
# so can reject null hypothesis H0 : b1 = 0
```


## (f) Do the results obtained in (c)–(e) contradict each other? Explain your answer.
```{r}
# NO, results do not contradict. 
# when predictors are highly collinear, 
# the change in response variable due to one predictor 
# can be masked/hidden by collinear predictor. 
```

## (g) Now suppose we obtain one additional observation, which was
unfortunately mis-measured. 
```{r}
x1 <- c(x1 , 0.1)
x2 <- c(x2 , 0.8)
y <- c(y, 6)
```

```{r}
cor(x1,x2)
```

```{r}
plot(x1,x2)
text(x=0.1, y=0.8, labels="new_data_point", pos=4, cex=0.7, col="red")
abline(x1,x2, col="blue")
legend(x="top",   # Coordinates of legend_box (x also accepts keywords)
       legend =c('abline'), # Vector with the name of each group
       fill,   # Creates boxes in the legend with the specified colors
       col = "blue", # Color of lines or symbols
       border = "black", # Fill box border color
       lwd=1,         # Line type and width
       bty = "o",        # Box type (bty = "n" removes the box)
       bg = par("bg"),    # Background color of the legend
       box.lwd = par("lwd"), # Legend box line width
       box.lty = par("lty"), # Legend box line type
       box.col = par("fg"),  # Legend box line color
       cex = 1,          # Legend size
       horiz = FALSE,     # Horizontal (TRUE) or vertical (FALSE) legend
       title = NULL      # Legend title
)
```
#### new data point seems to be an outlier.

## Re-fit the linear models from (c) to (e) using this new data. 
What effect does this new observation have on the each of the models?
```{r}
lm_4 <- lm(y~ x1+x2)
summary(lm_4)
```
```{r}
summary(lm_1)
```
#### for x1 & x2 predictor models; 
old lm_1 model showed x1 to be significant based on p-value<0.05 
and x2 not significant.
BUT, after adding new data point,
new lm_4 model shows x2 to be signficant based on p-value<0.05 
and x1 not significant.
Also RSE increased from old 1.056 to new 1.075.

```{r}
lm_5 <- lm(y~ x1)
summary(lm_5)
```

```{r}
summary(lm_2)

```
#### for x1 only predictor models; 
old lm_2 model showed RSE 1.055.
BUT, after adding new data point,
new lm_5 model shows RSE 1.111, so RSE increased.


```{r}
lm_6 <- lm(y~ x2)
summary(lm_6)
```
```{r}
summary(lm_3)
```
#### for x2 only predictor models; 
old lm_3 model showed RSE 1.072.
BUT, after adding new data point,
new lm_6 model shows RSE 1.074, so RSE increased.


## In each model, is this observation an outlier? 
A high-leverage point? 
Both? Explain your answers.
```{r}
par(mfrow = c(2, 2))
plot(lm_4)
```

```{r}
plot(predict(lm_4), rstudent(lm_4))
```

```{r warning=FALSE}
par(mfrow = c(1, 3))

#calculate leverage for each predictor data point per model
hats_lm_4 <- as.data.frame(hatvalues(lm_4)) 
hats_lm_5 <- as.data.frame(hatvalues(lm_5)) 
hats_lm_6 <- as.data.frame(hatvalues(lm_6)) 

plot(hats_lm_4, type = 'p')
title(main = 'lm_4, hat values plot')
text(x=0.4, y=1, labels="new_data", pos=2, cex=0.5, col="red")

plot(hats_lm_5, type = 'p')
title(main = 'lm_5, hat values plot')
text(x=0.05, y=1, labels="new_data", pos=1, cex=0.7, col="red")

plot(hats_lm_6, type = 'p')
title(main = 'lm_6, hat values plot')
text(x=0.1, y=1, labels="new_data", pos=3, cex=0.5, col="red")

```
#### as observed new data point is high leverage in models lm_4 and lm_6. 
```{r}
par(mfrow = c(1, 3))
plot(hatvalues(lm_4))
title(main = 'lm_4, hat values plot')
text(x=100, y=0.4, labels="new_data", pos=1, cex=0.5, col="red")

plot(hatvalues(lm_5))
title(main = 'lm_5, hat values plot')
text(x=27, y=0.045, labels="OLD_data", pos=1, cex=0.5, col="green")

plot(hatvalues(lm_6))
title(main = 'lm_6, hat values plot')
text(x=100, y=0.1, labels="new_data", pos=1, cex=0.5, col="red")

```
```{r}
which.max(hatvalues(lm_4))
```
### this is high leverage and outlier 
```{r}
which.max(hatvalues(lm_5))
```

```{r}
which.max(hatvalues(lm_6))
```
### this is high leverage and outlier 
###############################################################################################

